Climate indicators

### 2023-02-23 DRAFT_climate_indicators.Rmd
add noise earlier in workflow

@@

y1 <- x_summerMean_v2
y2 <- x_summerMean_v2 %>% st_set_dimensions("time", values = oldTime %m+% years(-1)) %>% mutate(Attribute_A = Attribute_A*rnorm(1,1,.3)) 
y3 <- x_summerMean_v2 %>% st_set_dimensions("time", values = oldTime %m+% years(-2)) %>% mutate(Attribute_A = Attribute_A*rnorm(1,1,.3))
y4 <- x_summerMean_v2 %>% st_set_dimensions("time", values = oldTime %m+% years(-3)) %>% mutate(Attribute_A = Attribute_A*rnorm(1,1,.3))
y5 <- x_summerMean_v2 %>% st_set_dimensions("time", values = oldTime %m+% years(-4)) %>% mutate(Attribute_A = Attribute_A*rnorm(1,1,.3))

@@

```{r}
y_ref <- y %>%
  filter(band==1) %>%
  st_apply(c("x", "y"), FUN =  median_sd)
```

@@

It's easier if median and sd are unique attributes, rather than levels of a dimension.

@@

The climate variable is the same as y1, y2, y~*n*~, ....
But to create some differences between the variable and the reference values, I will tweak these slightly, more or less divide them by two (plus some error). Actually, just leaving the reference layer and the variable layer perfectly correlated causes an error when dividing the dimension using st_apply.

```{r}

myRandom <- function(x) x*rnorm(1,1,.1)/2
myRandom(1)

y1v <- y1 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom)
y2v <- y2 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
y3v <- y3 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
y4v <- y4 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
y5v <- y5 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom)

#y_v <- c(y1v, y2v, y3v, y4v, y5v, y_ref)

y_v <- c(y1v, select(y_ref, reference_value))
y_v_merged <- merge(y_v)
dim(y_v_merged)
names(y_v_merged)
y_v_merged <- setNames(y_v_merged, "value")

@@

```{r}
myFUN <- function(a,b) (a/b)
system.time(
  temp <- st_apply(y_v_merged, 
                   c("x", "y"), 
                   #1:2,
                   FUN = myFUN
  )
)

@@

```{r}
tm_shape(temp)+
  tm_raster()+
  tm_layout(legend.outside = T)
```
This looks like perfect noise, and it is., because the variable and reference value are so well correlated.
If the reference value was static, it would look different.

@@

```{r}
system.time(
  temp <- st_apply(y_v_merged, 
                   c("x", "y"), 
                   #1:2,
                   FUN = myFUN
  )
)

@@

```{r}

y_scaled <- merge(y_v)
dim(y_scaled)
names(y_scaled)
```

```{r}
tm_shape(y_v_scaled)+
  tm_raster()
```

>>>>

Firt, let me cut the _band_ dimension.
```{r}
temp <- x_summerMean_v2 %>%
  filter(band==1) %>%
  adrop(drop=4) %>%
  aperm(c(2,3,1))
dim(temp)
```

@@

# Function to add random noise
myRandom <- function(x) x*rnorm(1,1,.1)

y1 <- temp
y2 <- temp %>% st_set_dimensions("time", values = oldTime %m+% years(-1)) %>% st_apply(c("x", "y", "time"), FUN = myRandom)
y3 <- temp %>% st_set_dimensions("time", values = oldTime %m+% years(-2)) %>% st_apply(c("x", "y", "time"), FUN = myRandom)
y4 <- temp %>% st_set_dimensions("time", values = oldTime %m+% years(-3)) %>% st_apply(c("x", "y", "time"), FUN = myRandom)
y5 <- temp %>% st_set_dimensions("time", values = oldTime %m+% years(-4)) %>% st_apply(c("x", "y", "time"), FUN = myRandom)

dim(y1)
dim(y2)

@@

```{r}
y_ref <- y %>%
  st_apply(c("x", "y"), FUN =  median_sd)
```

@@

It's perhaps easier if median and sd are unique attributes, rather than levels of a dimension.

@@

The climate variable is the same as y1, y2, y~*n*~, ....
But to create some differences between the variable and the reference values, I will tweak these slightly, more or less divide them by two.

```{r}
#y1v <- y1 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom)
#y2v <- y2 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
#y3v <- y3 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
#y4v <- y4 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
#y5v <- y5 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
#y_v <- c(y1v, y2v, y3v, y4v, y5v, y_ref)

y1v <- y1 %>% adrop() %>% mutate(Attribute_A = Attribute_A/2)
y_v <- c(y1v, select(y_ref, reference_value))
y_v_merged <- merge(y_v)
dim(y_v_merged)
names(y_v_merged)
y_v_merged <- setNames(y_v_merged, "value")

@@

```{r}
# A function for normalization
# I need to add a break point and lower reference value here as well.
# And I need to add truncation.
myFUN <- function(a,b) (a/b)
```

```{r}
system.time(
  temp <- st_apply(y_v_merged, 
                   c("x", "y"),
                   FUN = myFUN
  )
)

@@

```{r, fig.cap"The normalised indicator for year 1970"}
tm_shape(temp)+
  tm_raster()+
  tm_layout(legend.outside = T)
```
This looks like perfect noise, and it is, because the variable and reference value are so well correlated.
If the reference value was static, it would look different.

@@

```{r}
temp <- st_apply(y_v_merged, 
                   c("x", "y"), 
                   #1:2,
                   FUN = myFUN
)
```

@@
